Blueprint for efficient nuclear spin characterization in color center

### 2024-01-30 article.tex
text

\author{Majid Zahedian}
 \affiliation{3rd physical institute University of Stuttgart, Germany}
  %\affiliation{Max Planck School of Photonics, Germany}%Lines b%Lines break automatically or can be forced with \\

\author{Vadim Vorobyov}
\email[]{v.vorobyov@pi3.uni-stuttgart.de}
 \affiliation{3rd physical institute University of Stuttgart, Germany}  
% with \\

\author{J\"org Wrachtrup}
 \affiliation{3rd physical institute University of Stuttgart, Germany}%Lines break 
 \affiliation{Max Planck Institute for solid state physics, Stuttgart, Germany}%Lines break 
 %automatically or can be forced with \\ Third institution, the second for Charlie Author

@@

\begin{abstract}
Nuclear spins in solids is one of the most promising platforms for realisation of a scalable quantum hardware. They could be efficiently adressed at single site level by nearby single colour centers via the spin resonance. However, characterising each nuclear spin is quite cumbersome as the characterisation protocol may differ depending on the strength of the hyperfine coupling. The most adopted protocol is modified electron spin Hahn-echos such as CPMG and XY8 pulse sequences. In some cases, e.g. for spin 1/2, strongly coupled spins, and bispecies bath this protocol contains severe problems in applying for the nuclear spin bath characterisation. 
Here, we present a more straight forward method to obtain hyperfine interaction among nuclear spins and the electron spin. 
This method can be used for a variety of platform including emerging S=1/2 group IV defects in diamond (e.g. SiV, GeV, SnV, PbV) or Silicon vacancy in Silicon Carbide.
We summarize theoretical framework and adopt it from ESR for the use in colour centers with various spins, and numerically consider case of nuclear spin cluster, as compare performance of various protocols on this example via the Fisher information matrix. 
\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
%\tableofcontents
\section{Introduction and background}
Optically active defects in solids, color centers, has been applied for realising different quantum applications \cite{awschalom2018quantum}, such as a quantum network \cite{pompili2021realization}, quantum sensing \cite{degen2017quantum}, and quantum registers \cite{bradley2019ten}.

Each center contains an electron spin that can be controlled directly with microwaves, initialised and read out via optical excitation. 
The key element for most applications is that they could be coupled through the hyperfine coupling to the bath of numerous nuclear spins presenting a long living quantum memories and could present an optically accessible nuclear spin qubit register. 
To design the efficient quantum control for the memories, a precise knowledge of the full hamiltonian of the register needs to be known and characterised before. 
However, characterising the nuclear spin hyperfine coupling could be a challenging and a lengthy procedure \cite{taminuau27, taminiau50}. 
Traditionally, nuclear spin characterisation is done with Optically Detected Magnetic Resonance (ODMR) \cite{Dreau, Neumann, Eriks}, however the spectroscopic resolution is limited to the $1/T_2^\ast$ of the electron spin. 
Hahn Echo type sequences have been used to refocus the electron spin and enhance the coherence time of the electron spin \cite{childress2006coherent} thus improving the resolution to $1/T_2$. 
This method could be applied for the defects with a specific nuclear spins configuration \vvv{(which?)}. 
\vvv{In particular, the certain relation to the magnetic field and the coupling of the nuclear spins should be kept, (weakly coupled nuclear spins), as well as the spin number of the electron spin is affecting the observed behaviour}
Depending on the applied protocol for the nuclear spins characterization, the number of accessible nuclear spin can increase. Electron Spin Echo Envelop Modulation (ESEEM) types of sequences allow us to access the highest number of nuclear spins since the spectroscopy resolution is limited to the longitudinal relaxation time ($T_1$) of the system (see fig. \ref{fig:1}) \cite{Laraui, Vorobyov}. 
In particular we consider following common examples of nuclear spin registers ubiquitios in the applications. First, NV-like case with S=1, and weakly coupled nuclear spins, at high mangetic field such that $\gamma_n B \gg A_{zx}$. 
This case is very well studied in the works of \cite{Tam27, Tam2012, Tam2014, Wrachtrup, before}.
The ability to see narrow peaks with analytically predictable position and contrast allows one to solve the inverse problem of characterising the coupling of the nuclear spins to the electron spins. 
The second case (see in Fig. \ref{fig1b}) refers to a similar case but with a strongly coupled nuclear spin, which hinders the observation of a weakly coupled nuclear spins. 
This case in particular  is relevant for the case, where additional to the weakly coupled register a strongly coupled nuclear spin is used for e.g. repetitive  enhancment of the readout, via using the strongly coupled ancilla resolved in electron spin resonance spectra. 
The third case (Fig. 1c) is relevant for the S=1/2 case, where due to the lack of the offset in the average evolution of the nuclear spin, its dynamics weakly depends on the Azz and only in the second order on the Azx \cite{suckachev}. 
This leads to a poor distinguising ability of multiple nuclear spins and as a result lack of individual addressability of nuclear spins. 
Finally, in the case of other materials host platforms, a bispecies nuclear spin bath could be observed, where the resonance peaks referrring to different larmour frequencies of different species additionaly complicate the reconstruction of the hamiltonian, and for unambiguous reconstruction of the interactions, additionall measurements e.g. at different B field would be needed. 
In this work we argue that a correlation type sequence, based on 5p eseem could serve as a general framework as it promisses greate reconstruction potential for all above mentioned cases.

\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/drawing0.pdf}
		\caption{Comparing DD and 5pESEEM as two nuclear spin characterization method in different systems. a. An electron spin one system: in DD signal, each nuclear spin has a distinct resonance time up to first order of $A_{zz}$; the accuracy of the obtained hyperfine interaction is limited to $T_2$ of the electron spin. In the Fourier transform of the ESEEM signal, the two resonance frequency of each spin is present; the accuracy of the obtained hyperfine interaction is limited to $T_1$ of the electron spin b. An electron spin one system with two strongly coupled nuclear spins and many weakly coupled spins. No resonance time is detectable since the required condition for DD ($\omega_L \gg A_{zz}, A{zx}$) does not hold. However, resonant frequencies is still obtainable in the ESEEM signal c. An electron spin one half system: nuclear spins have more or less the same resonance time as it depends on the second order of hyperfine coupling. However, resonant frequency of each spins is distinguishable in the ESEEM signal. d. A system containing two different nuclear spin species: two nuclear spin bath can interfere in the DD signal, which makes the signal analysis more challenging. However, the species can be inferred from ESEEM signal since species have different Larmor frequencies.}
		\label{fig:1}
	\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%================= %THEORY
\section{Results}
\label{sec:theory}
%%%%%%%%%%%%%%%Í%%%%=================
We consider a central spin system of \textit{non-interacting} nuclear spins $I=1/2$ coupled to the central electron spin $\boldmath{S}$. 
The central electron spin is manipulated resonantly with the microwave pulses, which transfer the population between the two sublevels, denoted as $m_s=s_0$ and $m_s = s_1$. 
These two spin sublevels are separated, due to the Zeeman effect and/or zero field splitting, and form an two-level subsystem with energy splitting of $\omega_a$. 
The Hamiltonian of the system in the secular approximation and in the rotating frame of the applied microwave $\omega_{mw}$ can be written as:
\begin{align}
	H = \Delta S_z + \sum_k (\omega_L^{(k)}  + A_{zz}^{(k)} S_z) I_z^{(k)} + A_{zx}^{(k)} S_z I_x^{(k)}
	\label{eq:H}
\end{align}
Where $\Delta = \omega_a - \omega_{mw}$ is the detuning, $\omega_L^{(k)} =\gamma_n^{(k)} B$ is nuclear Rabi frequency with $\gamma_n^{(k)}$ nuclear spin gyromagnetic ratio and $B$ external magnetic field, $A_{zz}, A_{zx}$ parallel and perpendicular secular component of hyperfine tensor. The Hamiltonian is diagonal in the electron spin subspace, thus the hamiltonian for the nuclear spins could be rewritten in electron spin subdomains $s_i$ ($i=0,1$). The nuclear spin hamiltonian could be solved for eigenenergies, which determine the precession frequencies of $\omega_i = \sqrt{(\omega_L + s_i A_{zz})^2+ (s_i A_{zx})^2}$, $i=0,1$ along the eigenvector axes $\vec{n}_i =(\frac{s_i A_{zx}}{\omega_i}, 0, \frac{\omega_L + s_i A_{zz}} {\omega_i})$. It is assumed that the pulse duration $t_p$ are short enough ($\omega_L t_p \ll 1$) that the nuclear spin dynamics during this time is negligible. 
For completeness we start our analysis with the simplest characterization sequence, which is the electron spin ramsey sequence (free induction decay of electron) (see Appendix for derivation).
The Ramsey signal can be obtained as follows:
\begin{gather}
	\expval{\sigma_z}_{Ram} = \cos(\Delta \tau) \prod_{k=1}^{n} \Big[ \cos(\frac{\omega_0 \tau}{2}) \cos(\frac{\omega_1 \tau}{2})+ \notag \\
	(\vec{n_0}\cdot \vec{n_1})\sin(\frac{\omega_0 \tau}{2}) \sin(\frac{\omega_1 \tau}{2}) \Big]^{(k)}
	\textcolor{gray}{e^{-(\frac{\tau}{T_2^\ast})^m}}
\end{gather}	
We do not take any relaxation process into account and the gray exponential decay is put by hand to remember the relaxation time scale. Ramsey signal decay with electron spin $T_2^\ast$, so it can be used to observe the core of the defect. Ramsey sequence is sensitive to detuning since the electron spin is not refocused, however it is possible to sense nuclear spins with vanishing perpendicular hyperfine coupling.\\

To enhance the relaxation time and have access to more nuclei, one can add a $\pi$ pulse in between the Ramsey sequence and create Hahn-echo sequence and obtain the following signal:

@@

Without any nuclear spin, Hahn Echo sequence creates the electron spin echo signal with the envelope of stretched exponential decay with $T_2^{HE}$. 
However in the presence of nuclear spins, the electron spin echo envelope will be modulated due to interaction with nuclei. 
Hence, this sequence also called Electron Spin Echo Envelope Modulation or shortly ESEEM. 

@@

Hahn Echo sequence can be used for defects that contain a few strongly coupled nuclei. In order to differentiate the resonance frequency of each nuclear spin or make the oscillation sharper, one can to add more $\pi$ pulses and for the so-called Dynamical Decoupling sequence to sharpen the resonance conditions (neglecting decoherence terms):

@@

$\vec{n}_0 \cdot \vec{n}_1 = \frac{\omega_L^2+(s_0+s_1)A_{zz}+s_0 s_1 (A_{zz}^2+A_{zx}^2)}{\omega_0 \omega_1}$ and $N$ is the number of $\pi$ pulses that must be an even. Each nuclear 
If the magnetic field is high or all nuclear spins are weakly coupled such that $\omega_L \gg A_{zz}, A_{zx}$, the multiplication rule can be approximated by a summation rule, which makes the higher order resonance disappear:

@@

The parallel component of the hyperfine interaction can be obtained from the resonance time of each nuclear spin:

@@

@@

The perpendicular part of hyperfine interaction can be obtained from a fit to the Lorentzian dip with the width of $\frac{(s_1-s_0)\omega_L A_{zx}}{(2p+1)\pi\omega_0 \omega_1}$ or varying the number of pulses and fitting the minima position to the formula $\expval{\sigma_z}_{DD}^{(k)} \approx 1- 2 \sin^2\big(\frac{N}{2}\frac{(s_1-s_0)\omega_L A_{zx}}{\omega_0 \omega_1}\big) $. Although Dynamical Decoupling (DD) sequence is well understood and used to characterize nuclear spins, it is not applicable for any system. First, DD only works for defects with high magnetic field and low hyperfine coupling ($\omega_L \gg A_{zz}, A_{zx}$). Second, the signal is limited to electron spin $T_2$, which is difficult to make it equal to $T_1$. Third, the signal analysis is rather complicated and time-consuming. (Maybe AI solves this problem in the future?) one has to collect enough data to make sure no two nuclear spin overlap in the signal. Forth, this sequence does not work for spin one half system, e.g., group IV defect in diamond, since parallel component of hyperfine coupling cancels their effect in the resonance time up to the first and second order. Hence, the resonance time depends on the second order of perpendicular component. In this work, we take 23 nuclear spin characterized in paper \cite{abobeih2019atomic} as a realistic case. 
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/dd_compare2.pdf}
		\caption{Simulated DD signal with 64 $\pi$-pulses for the register reported \cite{abobeih2019atomic} assuming the electron spin is a. spin one b. spin one-half system. Simulated Fisher information for the case of the electron spin c. one d. one-half system.}
		\label{fig:3}
	\end{center}
\end{figure*}
In this work, we take the 23 nuclear spins cluster characterized in paper \cite{abobeih2019atomic} as a realistic case. 
First, we apply this to the dynamical decoupling sequence, and visualise the obtained results. 
Fig \ref{fig:3} compares DD signal for the same nuclear spin register but different defects in diamond. 
An intuitive description for the sensitivity of a sequence to the variation of hyperfine coupling could estimated knowing the width and sensitivity of the position of the resonances. 
First, we consider the case of NV center in diamond. 
The width of a dip $\delta \tau = \frac{A_{zx}}{\omega_L^2}$  and taking the derivative of the resonance time, we define the sensitivity of longitudinal hyperfine coupling as follows:
	\delta A_{zz} = \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zz}}}} = \frac{2}{\tau_k} \frac{A_{zx}}{\omega_L} \ge \frac{4}{T_2^{HE}} \frac{A_{zx}}{\omega_L}
If we assume the typical values of $T_2^{HE} = 100 $ $\mu$s and $\omega_L=500 $ kHz, a weakly coupled nuclear spin $A_{zx}=5$ kHz can be distinguished from another weakly coupled nuclear spin with $\delta A_{zz} = 400$ Hz. 
For $S=1/2$, e.g. for group IV defects in diamond, the resonance times is sensitive to the second order of $A_{zx}$ and no sensitivity to $A_{zz}$. 
Hence, even at large $\tau$ the resonances related to nuclear spins will not be separated well. 
We define the sensitivity for the spin one half system as follows:
	\delta A_{zx} = \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zx}}}} = \frac{4}{\tau_k} \ge \frac{8}{T_2^{HE}}
Assuming the typical values of $T_2^{HE} = 100 $ $\mu$s, a nuclear spin can be distinguished from another one if their transverse hyperfine coupling is separated by $\delta A_{zx} = 80$ kHz, which is quite inaccurate and $\approx 160$ times worse than for $S=1$.
To quantify the difference in sensitivity to estimating the hyperfine parameters, we perform calculation of the Fisher information matrix (FIM) for the 23 nuclear spin cluster, with 23 $A_{zz}$ and 23 $A_{zx}$ parameters (see Fig. \ref{fig:3}c,d). 
The fisher information matrix is estimated for the probability to measure state $|0\rangle$ and reads as: 
\begin{equation}
	F_{ij}(\textbf{A})=\sum_\tau \frac{\partial p(0, \tau, \textbf{A})}{\partial A_i}\frac{\partial p(0, \tau, \textbf{A})}{ \partial A_j} \frac{1}{p (1-p)}.
\end{equation}
For the typical case of NV center, with $S=1$, since all the nuclear spin resonances are clearly resolved, the Fisher information matrix takes a diagonal shape, revealing a low covariance between the various nuclear spin resonances. 
The crosstalk is existing between $A_{zz}$ and $A_{zx}$ of same nuclear spin, which is because the position of the peaks depends on both values. 
The fisher information gives a bound for the precision of parameter estimation, known as Cramér–Rao bound: 
\begin{equation}
	\delta A^2 \ge \frac{1}{F(A) N}
\end{equation}
For the presented protocol, the precision for various spins is limited by 0.09 Hz for the S=1, for the N = 4000. \\

The striking difference to the $S=1$ fisher information matrix, is the behaviour of the defect with $S=1/2$.

On the other hand, a sequence that is limited by $T_1$ instead of $T_2$ is based on 3 and 5-pulse ESEEM. The main principle of 5-pulse ESEEM \cite{kasumaj20085} is to create an entanglement (via e.g. Hahn-echo block) before the free evolution of the nuclear spins and after make a second correlating sequence. 
In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix, thus limitied by $T_1$ time. 
Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. 
The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.

>>>>

\author{Majid Zahedian}
 \affiliation{3rd Physical Institute University of Stuttgart, Germany}
 \affiliation{Institute of Quantum Electronics, ETH Z\"urich, Switzerland}%Lines b%Lines break automatically or can be forced with \\

\author{Vadim Vorobyov}
\email[]{v.vorobyov@pi3.uni-stuttgart.de}
 \affiliation{3rd Physical Institute University of Stuttgart, Germany}  
% with \\

\author{J\"org Wrachtrup}
 \affiliation{3rd Physical Institute University of Stuttgart, Germany}%Lines break 
 \affiliation{Max Planck Institute for solid state physics, Stuttgart, Germany}%Lines break 
 %automatically or can be forced with \\ Third institution, the second for Charlie Author

@@

\begin{abstract}
Nuclear spins in solids offer a promising avenue for developing scalable quantum hardware. Leveraging nearby single-color centers, these spins can be efficiently addressed at the single-site level through spin resonance.
However, characterising individual nuclear spin is quite cumbersome since the utility of characterisation protocols may differ depending on the strength of the hyperfine coupling, necessitating tailored characterization protocols.
While modified electron spin Hahn echoes like CPMG and XY8 pulse sequences are commonly employed, they encounter significant limitations in scenarios involving spin-1/2 systems, strongly coupled spins, or nuclear spin baths comprising distinct isotopes.
Here, we present a more straightforward approach for determining the hyperfine interactions among each nuclear and the electron spin. 
This method holds promise across diverse platforms, including emerging S=1/2 group IV defects in diamond (e.g., SiV, GeV, SnV, PbV) or silicon vacancies in silicon carbide. 
We provide a theoretical framework and adapt it for use with color centers exhibiting various spins. Through simulations conducted on nuclear spin clusters, we evaluate different protocols and compare their performance using the Fisher information matrix.
\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
%\tableofcontents
\section{Introduction and background}
Optically active defects in solids, known as color centers, have been utilized in various quantum applications \cite{awschalom2018quantum}, including quantum networks \cite{pompili2021realization}, quantum sensing \cite{degen2017quantum}, and quantum registers \cite{bradley2019ten}.

Each center contains an electron spin that can be directly controlled using microwaves, and it can be initialized and read out through optical excitation. 
The potential hyperfine interactions with a bath of numerous nuclear spins are crucial for diverse applications, as they offer long-lived quantum memories and enable the creation of an optically accessible nuclear spin qubit register.
To implement efficient quantum control for these memories, it is imperative to have precise knowledge of the full Hamiltonian governing the register.
However, characterizing the hyperfine coupling of nuclear spins can be both challenging and time-consuming \cite{abobeih2019atomic, van2023mapping}. 
Traditionally, nuclear spin characterization is accomplished through Optically Detected Magnetic Resonance (ODMR) \cite{Dreau, Neumann, Eriks}, but the spectroscopic resolution is constrained by the $1/T_2^\ast$ of the electron spin. 
To overcome this limitation, Hahn-Echo-type sequences have been employed to refocus the electron spin and extend its coherence time \cite{childress2006coherent}, thereby improving the resolution to $1/T_2$. 
This method could be applied to defects with a specific configuration of nuclear spins, particularly those with a certain relation to the magnetic field and the coupling of the nuclear spins, such as weakly coupled nuclear spins.
Also, the protocols also can only be applied to electron spins with a specific spin multiplicity
Depending on the applied protocol for the nuclear spins characterisation, the number of identifiable nuclear spin can increase.
Electron Spin Echo Envelop Modulation (ESEEM) types of sequences allow access to the highest number of nuclear spins, as the spectroscopy resolution is limited to the longitudinal relaxation time ($T_1$) of the system. (see fig. \ref{fig:1}) \cite{Laraui, Vorobyov}.
In particular, we consider the following common examples of nuclear spin registers that are ubiquitous in applications. First, the NV-like case with S=1 and weakly coupled nuclear spins, operating at high magnetic fields where $\gamma_n B \gg A_{zx}$ which has been extensively studied in the works of \cite{abobeih2019atomic, taminiau2012detection, Tam2014, Wrachtrup, before}.
The capability to observe narrow peaks with analytically predictable positions and contrasts enables the solution of the inverse problem concerning the characterization of the coupling between nuclear spins and electron spins. 
The second case (as depicted in Fig. \ref{fig:1}b) corresponds to a scenario similar to the first, but with the presence of a strongly coupled nuclear spin, which obstructs the observation of weakly coupled nuclear spins.
This case is particularly relevant in situations where, in addition to the weakly coupled register, a strongly coupled nuclear spin is utilized, for instance, for repetitive enhancement of the readout.
This enhancement is achieved by exploiting the strongly coupled ancilla, which is discernible in electron spin resonance spectra.
The third case (Fig. 1c) pertains to the S=1/2 scenario, where the lack of offset in the average evolution of the nuclear spin results in its dynamics weakly depending on Azz and only in the second order on $A_{zx}$ \cite{suckachev}.
Consequently, this leads to a limited ability to distinguish between multiple nuclear spins and, consequently, a lack of individual addressability of nuclear spins.
Lastly, in the case of other material host platforms, a bispecies nuclear spin bath may be observed, where resonance peaks corresponding to different Larmor frequencies of different species further complicate the reconstruction of the Hamiltonian.
For unambiguous reconstruction of the interactions, additional measurements, such as those at different magnetic field strengths, would be required.
In this work, we propose that a correlation-type sequence, based on five pulse ESEEM, could serve as a general framework as it holds promise for the reconstruction of the interactions in all of the aforementioned cases.

\begin{figure*}%[htb]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/drawing0.pdf}
		\caption{Comparing DD and five pulse ESEEM as two nuclear spin characterization methods in different systems. a. Electron spin one system: In the DD signal, each nuclear spin exhibits a distinct resonance time up to the first order of $A_{zz}$; the accuracy of the obtained hyperfine interaction is limited to $T_2$ of the electron spin. In the Fourier transform of the ESEEM signal, the two resonance frequencies of each spin are present; the accuracy of the obtained hyperfine interaction is limited to $T_1$ of the electron spin. b. Electron spin 1 system with two strongly coupled nuclear spins and many weakly coupled spins: No resonance time is detectable since the required condition for DD ($\omega_L \gg A_{zz}, A_{zx}$) does not hold. However, resonant frequencies are still obtainable in the ESEEM signal. c. Electron spin 1/2 system: Nuclear spins have more or less the same resonance time as it depends on the second order of hyperfine coupling. However, the resonant frequency of each spin is distinguishable in the ESEEM signal. d. System containing two different nuclear spin species: Two nuclear spin baths can interfere in the DD signal, making the signal analysis more challenging. However, the species can be inferred from the ESEEM signal since species have different Larmor frequencies.}
		\label{fig:1}
	\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%================= %THEORY
\section{Results}
\label{sec:theory}
%%%%%%%%%%%%%%%Í%%%%=================
We consider a central spin system of \textit{non-interacting} nuclear spins $I=1/2$ coupled to the central electron spin $\boldsymbol{S}$.
The central electron spin is manipulated resonantly with microwave pulses, which transfer the population between the two sublevels denoted as $m_s=s_0$ and $m_s = s_1$.
These two spin sublevels are separated due to the Zeeman effect and/or zero-field splitting, forming a two-level subsystem with an energy splitting of $\omega_a$.
The Hamiltonian of the system in the secular approximation and in the rotating frame of the applied microwave $\omega_{mw}$ can be written as:
\begin{align}
	H = \Delta S_z + \sum_k (\omega_L^{(k)}  + A_{zz}^{(k)} S_z) I_z^{(k)} + A_{zx}^{(k)} S_z I_x^{(k)}
	\label{eq:H}
\end{align}
Where $\Delta = \omega_a - \omega_{mw}$ is the detuning, $\omega_L^{(k)} =\gamma_n^{(k)} B$ is the nuclear Rabi frequency with $\gamma_n^{(k)}$ being the nuclear spin gyromagnetic ratio and $B$ the external magnetic field, and $A_{zz}$ and $A_{zx}$ are the parallel and perpendicular secular components of the hyperfine tensor, respectively.
The Hamiltonian is diagonal in the electron spin subspace, thus the Hamiltonian for the nuclear spins could be rewritten in the electron spin subdomains $s_i$ ($i=0,1$).
The nuclear spin Hamiltonian could be solved for eigenenergies, which determine the precession frequencies of $\omega_i = \sqrt{(\omega_L + s_i A_{zz})^2+ (s_i A_{zx})^2}$, $i=0,1$ along the eigenvector axes $\vec{n}i =(\frac{s_i A{zx}}{\omega_i}, 0, \frac{\omega_L + s_i A_{zz}} {\omega_i})$.
It is assumed that the pulse duration $t_p$ is short enough ($\omega_L t_p \ll 1$) that the nuclear spin dynamics during this time is negligible.
For completeness, we start our analysis with the simplest characterization sequence, which is the electron spin Ramsey sequence (free induction decay of the electron) (see the Appendix for derivation).
The Ramsey signal can be obtained as follows:
\begin{gather}
	\expval{\sigma_z}_{Ram} = \cos(\Delta \tau) \prod_{j=1}^{n} \Big[ \cos(\frac{\omega_0 \tau}{2}) \cos(\frac{\omega_1 \tau}{2})+ \notag \\
	(\vec{n_0}\cdot \vec{n_1})\sin(\frac{\omega_0 \tau}{2}) \sin(\frac{\omega_1 \tau}{2}) \Big]^{(j)}
	\textcolor{gray}{e^{-(\frac{\tau}{T_2^\ast})^m}}
\end{gather}	
The exponential decay is introduced to note the Ramsey signal decays with the electron spin $T_2^\ast$ since the formalism does not account for any relaxation processes. The Ramsey sequence is sensitive to detuning since the electron spin is not refocused; however, it is possible to sense nuclear spins with vanishing perpendicular hyperfine coupling.\\

To extend the relaxation time and access more nuclei, one can insert a $\pi$ pulse between the Ramsey sequence, creating a Hahn echo sequence, and obtain the following signal:

@@

Without any nuclear spin, the Hahn Echo sequence creates an electron spin echo signal with an envelope of stretched exponential decay with $T_2^{HE}$.
However, in the presence of nuclear spins, the electron spin echo envelope will be modulated due to interaction with the nuclei.
Hence, this sequence is also called Electron Spin Echo Envelope Modulation or simply ESEEM.

@@

The Hahn Echo sequence can be used for defects that contain a few strongly coupled nuclei.
In order to differentiate the resonance frequency of each nuclear spin or to sharpen the oscillations, one can add more $\pi$ pulses, creating the so-called Dynamical Decoupling sequence to separate the resonance condition for individual nuclear spin. For the even number of $\pi$ pulses, $N$, The DD signal can be obtained (neglecting decoherence terms):
\begin{gather} \label{eq:DD}

@@

Where $\vec{n}_0 \cdot \vec{n}_1 = \frac{\omega_L^2+(s_0+s_1)A_{zz}+s_0 s_1 (A_{zz}^2+A_{zx}^2)}{\omega_0 \omega_1}$.
At this point, it should be clear that the number of pulses can be used as an additional parameter to modify the modulation depth of each nuclear spin.
A detailed investigation of the obtained DD signal reveals that by sweeping the time interval between pulses, an exponential decay with the rate of $1/T_2$ is observed, except for some resonance times where the electron spin becomes entangled with a particular nuclear spin, resulting in a sharp drop in the signal.
This resonance time can be obtained assuming $\omega_L \gg A_{zz}, A_{zx}$ is satisfied: 
\begin{gather} \label{eq:resonance}
Experimental observation of the resonance times provides valuable information about the nuclear spins.
For instance, in the case of an NV center, the second-order term $\frac{A_{zx}^2}{\omega_L^2}$ can be neglected, allowing for the direct determination of the $A_{zz}$ hyperfine component from the resonances.
To proceed with the characterization, Eq. \ref{eq:DD} needs to be further simplified, requiring stronger assumptions.
Focusing on weakly coupled nuclear spins or high magnetic field situations, the multiplication rule can be approximated by a summation rule, neglecting the higher-order cross-resonance terms:

@@

Assuming non-overlapping resonance times (which does not hold true for spin 1/2 systems) and expanding the signal around the j-th nuclear spin resonance ($\tau = \tau_p^{(j)}+\delta \tau$) up to the first order of $\delta \tau$ results in approximating each drop with a Lorentzian function:

@@

	\expval{\sigma_z}_{DD} (\delta \tau ) \approx 1 -2\sin^2(\frac{N}{2}\frac{(s_1-s_0)A_{zx}}{\omega_L}) \frac{(\frac{(s_1-s_0)A_{zx}}{2\omega_L^2})^2}{{\delta \tau}^2 +(\frac{(s_1-s_0)A_{zx}}{2\omega_L^2})^2}

@@

This simplification allows for the determination of the parallel component of the hyperfine interaction for each nuclear spin, possibly in two ways. First, a Lorentzian with a width of $\frac{(s_1-s_0) A_{zx}}{\omega_L^2}$ can be fitted to a dip to identify $A_{zx}$. Second, by keeping track of the minima of a dip while varying the number of pulses, the periodic function $1 - 2 \sin^2\big(\frac{N}{2}\frac{(s_1-s_0) A_{zx}}{\omega_L}\big)$ can be fitted to obtain $A_{zx}$.

Although the Dynamical Decoupling (DD) sequence is well understood and used to characterize nuclear spins, it is not applicable to all systems.
First, DD only works for defects with high magnetic fields and low hyperfine coupling ($\omega_L \gg A_{zz}, A_{zx}$).
Second, the signal is limited to the electron spin $T_2$, which only gradually approaches $T_1$.
Third, signal analysis is rather complicated and time-consuming; one has to collect enough data to ensure that no two nuclear spins overlap in the signal.
Fourth, this sequence does not work for spin 1/2 systems, such as group IV defects in diamond, since the parallel component of the hyperfine coupling cancels their effect on the first order electron spin signal.
Hence, the resonance signal depends on the second order of the perpendicular hyperfine component.

In this work, we consider 23 nuclear spins characterized in paper \cite{abobeih2019atomic} as a realistic case.
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/dd_compare2.pdf}
		\caption{Simulated DD signal with 64 $\pi$-pulses for the register reported in  \cite{abobeih2019atomic} assuming the electron spin is a. spin 1 system b. spin 1/2 system. Simulated Fisher information matrix for the case of the electron spin c. spin 1 system d. spin 1/2 system.}
		\label{fig:3}
	\end{center}
\end{figure*}
In this work, we consider the 23-nuclear spin cluster characterized in paper \cite{abobeih2019atomic} as a realistic case for our comparison.
First, we apply this to the dynamical decoupling sequence and visualize the obtained results. Fig \ref{fig:3} compares the DD signal for the same nuclear spin register but different defects in diamond.
An intuitive description of the sensitivity of a sequence to the variation of hyperfine coupling could be estimated by considering the width and sensitivity of the position of the resonances.
First, we consider the case of an NV center in diamond. The width of a dip is $\delta \tau = \frac{A_{zx}}{\omega_L^2}$. Taking the derivative of the resonance time, we define the sensitivity of the longitudinal hyperfine coupling as follows, using Eq. \ref{eq:resonance} to simplify:
\begin{gather}
	\delta A_{zz} \equiv \frac{\delta \tau}{\abs{\frac{\partial \tau_p}{\partial A_{zz}}}} = \frac{2}{\tau_p} \frac{A_{zx}}{\omega_L} \ge \frac{4}{T_2^{HE}} \frac{A_{zx}}{\omega_L}
\end{gather}
If we assume typical values of $T_2^{HE} = 100$ $\mu$s and $\omega_L=500$ kHz, a weakly coupled nuclear spin $A_{zx}=5$ kHz can be distinguished from another weakly coupled nuclear spin with $\delta A_{zz} = 400$ Hz.
For $S=1/2$, such as group IV defects in diamond, the resonance times are sensitive to the second order of $A_{zx}$ and show no sensitivity to $A_{zz}$.
Hence, even at large $\tau$, the resonances related to nuclear spins will not be well separated.
We define the sensitivity for the spin one-half system as follows:
\begin{gather}
	\delta A_{zx} \equiv \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zx}}}} = \frac{4}{\tau_k} \ge \frac{8}{T_2^{HE}}
\end{gather}
Assuming typical values of $T_2^{HE} = 100$ $\mu$s, a nuclear spin can be distinguished from another one if their transverse hyperfine coupling is separated by $\delta A_{zx} = 80$ kHz, which is quite inaccurate and approximately 160 times worse than for $S=1$.
